Assessment and Optimization of Nowcast Measures for Trend Detection

### 2023-12-18 paper.tex
Update on Overleaf.

\begin{itemize}
    \item was ist mit dem Trendinghorizont, also welche Zeitinkremente sind zu nehmen. Da würde sich irgendwie ein Plot Maß gegen Inkrement anbieten. Dann würden sich für unterschiedliche Inkremente unterschiedliche Trendingcapabilities ergeben (gabe es nicht so einen schonmal bei irgendeinem Treffen?)
    \item Wahl des Rauschens; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen?
    \item Was ist mit anderen Verteilungen?
    \item Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
\end{itemize}

@@

\begin{itemize}
	\item Trending is an important feature of nowcasts to be identify changes in the situation before the actual quantity can be measured
	\item What scores are currently used?
	\item Examples: Covid, BIP, Medicine, ... 
\end{itemize}

@@

Notation
\begin{itemize}
  \item Let $y_1, \ldots, y_T$ be the values of the nowcasted quantity.
  \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps. If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an updated nowcast for time $t-1$. (Move this point to a place after error model, as the error model can account for the error made at this point)
  \item Then, $\xdiff_t^k$ can be seen as point forecast of nowcaster $k$ for $\ydiff_t$. 
\end{itemize}

@@

Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$

@@

\begin{itemize}
  \item Accuracy: \begin{equation}
  	m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-1}
\end{equation}
	\item Capability of detecting a positive trend (probability of detection)
	\begin{equation}
  		m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}
	\end{equation}
	\item False alarm rate \hl{Name does not fit} \begin{equation}
  m_{\text{far}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}
\end{equation}
maybe use \enquote{positive} measure instead, i.e.,
\begin{equation}
  m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}  
\end{equation}
	\item same measures for negative direction:
	\begin{align}
		m_{\text{ny}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} \\
		m_{\text{nx}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}
	\end{align}
 \item Bezug auf Confusionsmatrix 
\end{itemize}

@@

% \begin{itemize}
%   \item Quadratic/Brier Score for $k$ possible outcomes \hl{Erst mal rauslassen; wäre anderes setting: probabilistischer Nowcast}: \begin{equation}
%   S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
% \end{equation}
% for dichotomous outcomes ($p_1 = \hat{P}(\ydiff_t > 0)$: probability for $\ydiff_t > 0$):
% \begin{equation}
%   S(p_1, \ydiff_t) = p_1 \ind{\ydiff_t > 0} + (1 - p_1) (1 - \ind{\ydiff_t > 0}) + p_1^2 + (1-p_1)^2 - 1
% \end{equation}
% \end{itemize}
\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
\begin{itemize}
  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$:
  	\begin{align}\label{additive error decomposition}
  		\ydiff_t &= \ydifft + \varepsilon_t^y \\
  		\xdiff_t &= \xdifft + \varepsilon_t^x
	\end{align}
 \item \hl{Wouldn't it be more natural to model the original time series with a noise term?} (Qualitatively, it would yield the same results)
 \begin{align}
     y_t &= y_t^\star + \epsilon_t^y \\
     x_t &= x_t^\star + \epsilon_t^x \\
     \ydiff_t &= y_t^\star + \epsilon_t^y - (y_{t-1}^\star + \epsilon_{t-1}^y) = y_t^\star - y_{t-1}^\star + \underbrace{\epsilon_t^y - \epsilon_{t-1}^y}_{\sim N(0, \sqrt{2} \sigma_y)} \\
     \xdiff_t &= x_t^\star + \epsilon_t^x - (x_{t-1}^\star + \epsilon_{t-1}^x) = x_t^\star - x_{t-1}^\star + \underbrace{\epsilon_t^x - \epsilon_{t-1}^x}_{\sim N(0, \sqrt{2} \sigma_x)}
 \end{align}
 Advantage of the approach: For different lags, no new standard deviation has to be estimated.\\
 Disadvantage: Theoretically, there is serial correlation of the errors.
\item  Replace indicators in the above formulations by the corresponding probabilities:
\begin{align}
		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-1}  \\
   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
\end{align} 
\item Simple error model:
  \begin{align}
	  \varepsilon_Y \sim N(0, \sigma_Y) \\
	  \varepsilon_X \sim N(0, \sigma_X)
  \end{align}
  yields, e.g.,
  	\begin{equation}
  		\mprob_{\text{acc}} = \frac{1}{T-1} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
	\end{equation}
	where $\Phi$ denotes a (possibly multivariate) normal distribution
\end{itemize}
\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
\end{itemize}

@@

% \begin{figure}
%   \centering
% \includegraphics[scale=0.5]{plots/std_estimate.png}
%   \caption{Estimation of std with different methods.}
%   \label{fig:estimation std}
% \end{figure}

\begin{figure}
  \centering
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
    \end{subfigure}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
    \end{subfigure}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
    \end{subfigure}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
    \end{subfigure}
  \caption{Estimation of std with different methods.}
  \label{fig:estimation std}
\end{figure}

>>>>

\begin{itemize}
    \item was ist mit dem Trendinghorizont, also welche Zeitinkremente sind zu nehmen. Da würde sich irgendwie ein Plot Maß gegen Inkrement anbieten. Dann würden sich für unterschiedliche Inkremente unterschiedliche Trendingcapabilities ergeben (gabe es nicht so einen schonmal bei irgendeinem Treffen?)
    \item Wahl des Rauschens; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen?
    \item Was ist mit anderen Verteilungen?
    \item Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
    \item Messen die Maße das richtige? Sollte man andere Maße nutzen? Braucht man Korrekturterm?
    \item Indikatoren oder Wahrscheinlichkeiten?
    \item Wahrscheinlichkeit von der Pike herleiten
\end{itemize}

@@

\begin{itemize}
	\item Trending is an important feature of nowcasts to be identify changes in the situation before the actual quantity can be measured
	\item What scores are currently used?
	\item Examples: Covid, BIP, Medicine, ... 
 \item nowcasting schön aufschreiben, dass auch für Nicht-Nowcaster klar wird, warum Nowcasting speziell ist
\end{itemize}

@@

Notation
\begin{itemize}
  \item Let $y_1, \ldots, y_T$ be the values of the nowcasted quantity.
  \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps. If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an updated nowcast for time $t-1$. (Move this point to a place after error model, as the error model can account for the error made at this point) \todo{Kaligraphisches $x$, um vorzubereiten, dass es anderes sein könnte}
  \item Then, $\xdiff_t^k$ can be seen as point forecast of nowcaster $k$ for $\ydiff_t$. 
\end{itemize}

@@

Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$ (Problem: mit kleinen Fehlern; Hinweis Exclusion area)

@@

\begin{itemize}
  \item Accuracy: \begin{equation}
  	m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-w}
\end{equation}
	\item Capability of detecting a positive trend (probability of detection)
	\begin{equation}
  		m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}
	\end{equation}
	\item False alarm rate \hl{Name does not fit} \begin{equation}
  m_{\text{far}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}
\end{equation}
maybe use \enquote{positive} measure instead, i.e.,
\begin{equation}
  m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}  
\end{equation}
	\item same measures for negative direction:
	\begin{align}
		m_{\text{ny}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} \\
		m_{\text{nx}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}
	\end{align}
 \item Bezug auf Confusionsmatrix 
\end{itemize}

@@

% \begin{itemize}
%   \item Quadratic/Brier Score for $k$ possible outcomes \hl{Erst mal rauslassen; wäre anderes setting: probabilistischer Nowcast}: \begin{equation}
%   S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
% \end{equation}
% for dichotomous outcomes ($p_1 = \hat{P}(\ydiff_t > 0)$: probability for $\ydiff_t > 0$):
% \begin{equation}
%   S(p_1, \ydiff_t) = p_1 \ind{\ydiff_t > 0} + (1 - p_1) (1 - \ind{\ydiff_t > 0}) + p_1^2 + (1-p_1)^2 - 1
% \end{equation}
% \end{itemize}
\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
\begin{itemize}
  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
  	\begin{align}\label{additive error decomposition}
  		\ydiff_t &= \ydifft + \varepsilon_t^y \\
  		\xdiff_t &= \xdifft + \varepsilon_t^x
	\end{align}
 \item \hl{Wouldn't it be more natural to model the original time series with a noise term?} (Qualitatively, it would yield the same results)
 \begin{align}
     y_t &= y_t^\star + \epsilon_t^y \\
     x_t &= x_t^\star + \epsilon_t^x \\
     \ydiff_t &= y_t^\star + \epsilon_t^y - (y_{t-1}^\star + \epsilon_{t-1}^y) = y_t^\star - y_{t-1}^\star + \underbrace{\epsilon_t^y - \epsilon_{t-1}^y}_{\sim N(0, \sqrt{2} \sigma_y)} \\
     \xdiff_t &= x_t^\star + \epsilon_t^x - (x_{t-1}^\star + \epsilon_{t-1}^x) = x_t^\star - x_{t-1}^\star + \underbrace{\epsilon_t^x - \epsilon_{t-1}^x}_{\sim N(0, \sqrt{2} \sigma_x)}
 \end{align}
 Advantage of the approach: For different lags, no new standard deviation has to be estimated.\\
 Disadvantage: Theoretically, there is serial correlation of the errors. (Wie behandle ich die Autokorrelation; wie rausrechnen)
\item  Replace indicators in the above formulations by the corresponding probabilities:
\begin{align}
		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-w}  \\
   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
\end{align} 
\item Simple error model:
  \begin{align}
	  \varepsilon_Y \sim N(0, \sigma_Y) \\
	  \varepsilon_X \sim N(0, \sigma_X)
  \end{align}
  yields, e.g.,
  	\begin{equation}
  		\mprob_{\text{acc}} = \frac{1}{T-w} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
	\end{equation}
	where $\Phi$ denotes a (possibly multivariate) normal distribution
\end{itemize}
\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
\end{itemize}

@@

%\begin{figure}
%  \centering
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
%    \end{subfigure}
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
%    \end{subfigure}
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
%    \end{subfigure}
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
%    \end{subfigure}
%  \caption{Estimation of std with different methods.}
%  \label{fig:estimation std}
%\end{figure}%

\begin{figure}
  \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_25.pdf}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_50.pdf}
    \end{subfigure}

    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_100.pdf}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_1000.pdf}
    \end{subfigure}

  \caption{Estimation of std with different methods.}
  \label{fig:estimation std}
\end{figure}
